High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-07-20 inst/manuscript/manuscript.Rmd
Abstract redrafted

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: no
spacing: doublespacing
footerdate: no
note: This pre-print not yet undergone peer review. It has been submitted to the \emph{Frontiers in Artificial Intelligence}. This version will be updated as it is revised, and the final published version will be accessible through its DOI link.   
abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands. Such data are often unavailable for regions dominated by smallholder agricultural systems, which are particularly challenging for remote sensing. To overcome these challenges, we designed a system to minimize several sources of error that arise when mapping smallholder croplands. To overcome errors caused by mismatches between image resolution and cropland scales, as well as persistent cloud cover, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year. To reduce errors that occur when training classifiers, we built a labelling platform that rigorously assesses label accuracy, and creates more accurate consensus labels that train a Random Forests model. The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which is used in a segmentation process to delineate individual field boundaries. We applied this system to map Ghana’s croplands for the year 2018. We divided Ghana into 16 mapping regions (12,160-23,535 km$^2$), training separate models for each using a total of 6,299 labels, plus 1,600 for validation. Using an independent map reference sample (n=1,207), we found that overall accuracies of the resulting cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%. Croplands covered 16.1-23.2% of the mapped area, comprising 1,131,146 total fields with an average size of 3.92 ha. Estimates based on the map reference sample indicate the cropland percentage is 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on the map used to estimate the standard error. Using the labellers’ digitized field boundaries to estimate biases in field boundary statistics, we calculated an adjusted mean field size of 1.73 ha and total field count of 1,662,281. Although the cropland class contained substantial errors, the system was effective in mitigating error and quantifying resulting performance gains. By minimizing training errors, consensus labelling improved the model’s F1 scores by up to 25%, while 3 iterations of active learning increased the F1 score by 9.1%, on average, which was 2.3% higher than training models with randomly selected labels. Map accuracy can be improved by replacing Random Forests with a convolutional neural network. These results demonstrate a readily adapted, transferrable framework for developing high resolution, annual, nation-scale maps that provide important details about smallholder-dominated croplands.

@@

## Broader applications
This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using an approach that follows recommended best practices for training and assessing machine learning models [@ElmesAccountingtrainingdata2020]. This approach can be readily updated to integrate improvements, such as newer machine learning models. Beyond providing valuable insights into field characteristics, field boundary maps can help improve remote estimation of crop areas and yield [e.g. @EstesProjectedclimateimpacts2013], and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and farm size [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985]. Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of Africa.

>>>>

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: no
spacing: doublespacing
footerdate: no
note: This pre-print not yet undergone peer review. It has been submitted to the \emph{Frontiers in Artificial Intelligence}. This version will be updated as it is revised, and the final published version will be accessible through its DOI link.   
abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is key to understanding critical questions related to food security and other pressing socioeconomic and environmental concerns. However, accurately mapping these systems is difficult because of two key problems: 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers. We developed a cropland mapping approach designed to address these two problems, which we used to map Ghana’s croplands (excluding those under tree crops) for the year 2018.  To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering the 2018 agricultural calendar. To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a machine learning classifier (Random Forests) using an active learning approach, in which we used the uncertainty in model predictions to select more informative training samples. Minimizing label errors improved model F1 scores by up to 25%, while active learning increased the F1 score by an average of 9.1% between first and last training iterations, which was 2.3% higher than models trained with randomly selected labels. We used the resulting <4 m map of cropland probabilities within a final segmentation step to delineate crop field boundaries. Using an independent map reference sample (n=1,207), we found that overall accuracies of the cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%. Relying on the map reference sample to generate an unbiased area estimate from the field boundary map, we calculate that cropland covers 17.1 (15.4-18.9%). Using the labellers’ digitized field boundaries to estimate biases in the size and number of segments in the field boundary map, we estimate that there are 1,662,281 crop fields in Ghana with an average size of 1.73 ha. Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively minimize the errors inherent in remote sensing of smallholder-dominated agriculture.

@@

## Broader applications
This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using an approach that follows recommended best practices for training and assessing machine learning models [@ElmesAccountingtrainingdata2020]. This approach can be readily updated to integrate improvements, such as newer machine learning models. Beyond providing valuable insights into field characteristics, field boundary maps can help improve remote estimation of crop areas and yield [e.g. @EstesProjectedclimateimpacts2013], and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and farm size [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985]. Such maps will be important for understanding the rapid agricultural change that is currently occurring in Africa.
